---
output:
    word_document:
keep_tex: true
citation_package: natbib
fig_caption: no
title: >- 
  Supplement for: Using Digital Surveillance Tools for 
  Near Real-Time Mapping of the Risk of International Infectious Disease
  Spread: A Case Study
author: "Sangeeta Bhatia, Britta Lassmann, Emily Cohn, Moritz Kraemer, Mark Herringer, John Brownstein, Larry Madoff, Anne Cori, Pierre Nouvellet"
affiliation: Imperial College London
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: double
biblio-style: apsr
endnote: no
number_sections: yes
toc: yes
header-includes:
  - \usepackage{caption}
  - \captionsetup[figure]{labelformat=empty}
---

```{r SI-1, echo = FALSE}
source("si_figcaptions.R")
```

# Overview {-}

In this supplement, we present the details of the pre-processing of
ProMED and HealthMap feeds `r dataclean_section`, 
the model results using ProMED, HealthMap
and WHO data and the impact of the datasources and model parameters on
the performance of the model. 
We varied the length of the time window used for model fitting (see
Methods for details). `r pm_section`, `r hm_section` and
`r who_section` present the forecasts using ProMED, HealthMap and WHO
data respectively using different time windows (2, 4 and 6 weeks) 
and forecast horizons (4, 6 and 8 weeks). The time window did not have
a strong influence on the model performance 
(`r si_fig_nums(name = "pm_model_perf_all_windows", display = "c")`). 
Similarly, the data used also did not
influence the model performance 
(`r si_fig_nums(name = "model_perf_all_ds", display = "c")`). 
We explored the impact of alternate priors for the gravity model parameter 
on the results from the model. The results of this sensitivity analysis
are presented in `r gamma_ul_section`.

# Data cleaning

```{r SI-2, echo = FALSE, fig.cap = pm_data_clean, fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/27062019_promed_loglinear_Sierra Leone.pdf"))
```

## Correlation between R estimates 

```{r SI-3, echo = FALSE, message = FALSE, warning = FALSE}
coeffs <- readr::read_csv(
  file = here::here(
    "data/output",
    "Rt_median_who_pm_hm_correlation.csv"
  )
)

coeffs <- dplyr::select(coeffs, country:HealthMap_ProMED)
```


```{r SI-no, echo = FALSE, message = FALSE, warning = FALSE, eval=TRUE}
ft <- flextable::flextable(coeffs)
ft <- flextable::set_caption(ft, corr_coeffs)
ft <- flextable::autofit(ft)
ft
```

```{r SI-4, echo = FALSE, message = FALSE, warning = FALSE, eval=FALSE}
knitr::kable(coeffs, "latex", booktabs = TRUE)
```
# Forecasts using ProMED data
## Calibration window of 2 weeks
### Forecast horizon 6 weeks

```{r SI-5, echo = FALSE, fig.cap = pm_14_42_cap, fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/projections-viz-fixed-country-15-1.ProMED_14_42.pdf"))
```

### Forecast horizon 8 weeks

```{r SI-6, echo = FALSE, fig.cap = proj_figcaps[[1]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/projections-viz-fixed-country-15-1.ProMED_14_56.pdf"))
```

## Calibration window of 4 weeks
### Forecast horizon 4 weeks

```{r SI-7, echo = FALSE, fig.cap = proj_figcaps[[2]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/projections-viz-fixed-country-15-1.ProMED_28_28.pdf"))
```


### Forecast horizon 6 weeks


```{r SI-8, echo = FALSE, fig.cap = proj_figcaps[[3]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/projections-viz-fixed-country-15-1.ProMED_28_42.pdf"))
```


### Forecast horizon 8 weeks


```{r SI-9, echo = FALSE, fig.cap = proj_figcaps[[4]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/projections-viz-fixed-country-15-1.ProMED_28_56.pdf"))
```

## Calibration window of 6 weeks
### Forecast horizon 4 weeks

```{r SI-10, echo = FALSE, fig.cap = proj_figcaps[[5]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/projections-viz-fixed-country-15-1.ProMED_42_28.pdf"))
```

### Forecast horizon 6 weeks


```{r SI-11, echo = FALSE, fig.cap = proj_figcaps[[6]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/projections-viz-fixed-country-15-1.ProMED_42_42.pdf"))
```

### Forecast horizon 8 weeks



```{r SI-12, echo = FALSE, fig.cap = proj_figcaps[[7]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/projections-viz-fixed-country-15-1.ProMED_42_56.pdf"))
```


# Forecasts using HealthMap data
## Calibration window of 2 weeks
### Forecast horizon 4 weeks

```{r SI-13, echo = FALSE, fig.cap = proj_figcaps[[8]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/HealthMap/projections-viz-fixed-country-15-1.HealthMap_14_28.pdf"))
```

### Forecast horizon 6 weeks


```{r SI-14, echo = FALSE, fig.cap = proj_figcaps[[9]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/HealthMap/projections-viz-fixed-country-15-1.HealthMap_14_42.pdf"))
```

### Forecast horizon 8 weeks


```{r SI-15, echo = FALSE, fig.cap = proj_figcaps[[10]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/HealthMap/projections-viz-fixed-country-15-1.HealthMap_14_56.pdf"))
```

## Calibration window of 4 weeks
### Forecast horizon 4 weeks


```{r SI-16, echo = FALSE, fig.cap = proj_figcaps[[11]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/HealthMap/projections-viz-fixed-country-15-1.HealthMap_28_28.pdf"))
```

### Forecast horizon 6 weeks


```{r SI-17, echo = FALSE, fig.cap = proj_figcaps[[12]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/HealthMap/projections-viz-fixed-country-15-1.HealthMap_28_42.pdf"))
```


### Forecast horizon 8 weeks


```{r SI-18, echo = FALSE, fig.cap = proj_figcaps[[13]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/HealthMap/projections-viz-fixed-country-15-1.HealthMap_28_56.pdf"))
```


## Calibration window of 6 weeks
### Forecast horizon 4 weeks

```{r SI-19, echo = FALSE, fig.cap = proj_figcaps[[14]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/HealthMap/projections-viz-fixed-country-15-1.HealthMap_42_28.pdf"))
```

### Forecast horizon 6 weeks

```{r SI-20, echo = FALSE, fig.cap = proj_figcaps[[15]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/HealthMap/projections-viz-fixed-country-15-1.HealthMap_42_42.pdf"))
```

### Forecast horizon 8 weeks

```{r SI-21, echo = FALSE, fig.cap = proj_figcaps[[16]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/HealthMap/projections-viz-fixed-country-15-1.HealthMap_42_56.pdf"))
```


# Forecasts using WHO data
## Calibration window of 2 weeks
### Forecast horizon 4 weeks

```{r SI-22, echo = FALSE, fig.cap = proj_figcaps[[17]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/WHO/projections-viz-fixed-country-15-1.WHO_14_28.pdf"))
```

### Forecast horizon 6 weeks


```{r SI-23, echo = FALSE, fig.cap = proj_figcaps[[18]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/WHO/projections-viz-fixed-country-15-1.WHO_14_42.pdf"))
```

### Forecast horizon 8 weeks

```{r SI-24, echo = FALSE, fig.cap = proj_figcaps[[19]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/WHO/projections-viz-fixed-country-15-1.WHO_14_56.pdf"))
```

## Calibration window of 4 weeks
### Forecast horizon 4 weeks

```{r SI-25, echo = FALSE, fig.cap = proj_figcaps[[20]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/WHO/projections-viz-fixed-country-15-1.WHO_28_28.pdf"))
```

### Forecast horizon 6 weeks


```{r SI-26, echo = FALSE, fig.cap = proj_figcaps[[21]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/WHO/projections-viz-fixed-country-15-1.WHO_28_42.pdf"))
```

### Forecast horizon 8 weeks

```{r SI-27, echo = FALSE, fig.cap = proj_figcaps[[22]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/WHO/projections-viz-fixed-country-15-1.WHO_28_56.pdf"))
```

## Calibration window of 6 weeks
### Forecast horizon 4 weeks

```{r SI-28, echo = FALSE, fig.cap = proj_figcaps[[23]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/WHO/projections-viz-fixed-country-15-1.WHO_42_28.pdf"))
```

### Forecast horizon 6 weeks


```{r SI-29, echo = FALSE, fig.cap = proj_figcaps[[24]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/WHO/projections-viz-fixed-country-15-1.WHO_42_42.pdf"))
```

### Forecast horizon 8 weeks


```{r SI-30, echo = FALSE, fig.cap = proj_figcaps[[25]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/WHO/projections-viz-fixed-country-15-1.WHO_42_56.pdf"))
```

# Impact of time window on model performance

```{r SI-31, echo = FALSE, fig.align = "center", fig.cap = pm_model_perf_all_windows}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/forecasts-assess-by-time-window-13-1.ProMED_28_assess_by_twindow.pdf"))
```


# Impact of datasource on model performance


```{r SI-32, echo = FALSE,  fig.align = "center", fig.cap = model_perf_all_ds}
knitr::include_graphics(here::here("ms-figures/si-figures/other/forecasts-assess-by-datasource-6-1.pdf"))
```

# Sensitivity Analysis

For the results presented in the main text, we choose an uninformative
uniform prior for the parameter $\gamma$ with an upper bound 2. We
then fitted the model with a uniform prior for $\gamma$ allowing it to
vary from 1 to 10. In this section we present the results from the
sensitivity analyses. Although the analysis was carried out for the
three data sources (ProMED, HealthMap and WHO), for brevity we present
results using ProMED data only. 


## Mobility Model Parameters

```{r SI-33, echo = FALSE, fig.cap = pm10_params, fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/sensitivity_analyses/gravity_model_pars_14.pdf"))
```

## Forecasts using ProMED data
### Calibration window of 2 weeks
#### Forecast horizon 4 weeks

```{r SI-34, echo = FALSE, fig.cap = pm_figcaps[[1]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/sensitivity_analyses/projections-viz-fixed-country-15-1.ProMED_14_28.pdf"))
```


#### Forecast horizon 6 weeks

```{r SI-35, echo = FALSE, fig.cap = pm_figcaps[[2]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/sensitivity_analyses/projections-viz-fixed-country-15-1.ProMED_14_42.pdf"))
```



#### Forecast horizon 8 weeks

```{r SI-36, echo = FALSE, fig.cap = pm_figcaps[[3]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/sensitivity_analyses/projections-viz-fixed-country-15-1.ProMED_14_56.pdf"))
```

### Calibration window of 4 weeks
#### Forecast horizon 4 weeks

```{r SI-37, echo = FALSE, fig.cap = pm_figcaps[[4]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/sensitivity_analyses/projections-viz-fixed-country-15-1.ProMED_28_28.pdf"))
```

#### Forecast horizon 6 weeks


```{r SI-38, echo = FALSE, fig.cap = pm_figcaps[[5]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/sensitivity_analyses/projections-viz-fixed-country-15-1.ProMED_28_42.pdf"))
```

#### Forecast horizon 8 weeks


```{r SI-39, echo = FALSE, fig.cap = pm_figcaps[[6]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/sensitivity_analyses/projections-viz-fixed-country-15-1.ProMED_28_56.pdf"))
```

### Calibration window of 6 weeks
#### Forecast horizon 4 weeks

```{r SI-40, echo = FALSE, fig.cap = pm_figcaps[[7]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/sensitivity_analyses/projections-viz-fixed-country-15-1.ProMED_42_28.pdf"))
```

#### Forecast horizon 6 weeks


```{r SI-41, echo = FALSE, fig.cap = pm_figcaps[[8]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/sensitivity_analyses/projections-viz-fixed-country-15-1.ProMED_42_42.pdf"))
```

#### Forecast horizon 8 weeks

```{r SI-42, echo = FALSE, fig.cap = pm_figcaps[[9]], fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/sensitivity_analyses/projections-viz-fixed-country-15-1.ProMED_42_56.pdf"))
```

## Model performance with alternate priors for mobility model parameter

```{r SI-43, echo = FALSE, fig.cap = model_perf_gamma, fig.align = "center"}
knitr::include_graphics(here::here("ms-figures/si-figures/ProMED/sensitivity_analyses/forecasts_assess_by_gamma_ul.pdf"))
```
