---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
title: Cleaning up the feed from HealthMap/ProMED
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "`r Sys.Date()`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  infile: data/raw/promed_2014-2016-renamed.csv
  outfile: 19032019_promed_loglinear
  ofinterest: data/processed/all_african_centroids.csv
---


```{r data-cleanup-1, include=FALSE}
knitr::opts_chunk$set(fig.width = 12, fig.height = 6, echo = FALSE,
                      warning = FALSE, message = FALSE,
                      fig.path = "figures/")

```


```{r setup}
library(ggplot2)
library(ggthemes)
library(scales)
library(dplyr)
library(mRIIDS)
```


```{r data-cleanup-2 }

species   <- "Humans"
disease   <- "Ebola"
case.type <- "scc"
wafrica <- c("Sierra Leone",
              "Liberia",
              "Guinea",
             "Senegal",
             "Ghana",
             "Mali",
             "Nigeria")


```



Visualising the raw data.



```{r data-cleanup-3}

feed <- here::here(params$infile) %>%
             read.csv(stringsAsFactors = FALSE) %>%
    filter(Species == species &
           Disease == disease &
           Country %in% wafrica)

feed <- janitor::clean_names(feed)
## Janitor renames healthmap to health_map. Fix.
feed <- dplyr::rename(feed,
                      healthmap_alert_id = health_map_alert_id)


```

Strip the date of the time stamp
```{r data-cleanup-4}
feed <- tidyr::separate(feed,
                        issue_date,
                        sep = " ",
                        into = c("issue_date", "time"),
                        remove = TRUE
                        )
```



Split the data by country.

```{r data-cleanup-5 }
by.location <- feed %>%
    split(.$country) 

```

### Raw data by case category

```{r data-cleanup-6 }
raw_p <- purrr::map(by.location, function(x) {
    x <- select(x, issue_date, sc, sd, cc, cd, country) %>%
        tidyr::gather(case_type,
                      count, -c(issue_date, country)) 
    x$issue_date <- as.Date(x$issue_date,
                            format = "%m/%d/%y")
    p <- ggplot(x ,
                aes(issue_date, count, col = case_type)) +
        geom_point()
    p <- p + theme_classic()
    p <- p + xlab("") + ylab("Cases")
    p <- p + theme(panel.border=element_blank(), axis.line = element_line())
    p <- p + theme(axis.text.x = element_text(angle = 80, hjust = 1))
    p <- p + theme(legend.title = element_blank())
    p <- p + xlab("")
    p <- p + scale_x_date(date_labels =  "%b %Y")
    p <- p + ggtitle(x$country[1])
    p <- p + scale_color_discrete(
                         breaks = c("sc", "sd", "cc", "cd"),
                         labels = c("SC", "SD", "CC", "CD"))
    p
})

```

## 

```{r data-cleanup-7 }
ggpubr::ggarrange(plotlist = raw_p, common.legend = TRUE) %>%
    ggpubr::ggexport(filename = here::here("data/output/figures",
                                           paste0(params$outfile,
                                                  "_raw.png")),
                     res = 144,
                     width = 960,
                     height = 540)
```
# Data clean-up

Extract the total case count as a sum of suspected and
confirmed cases.


```{r data-cleanup-8 }

cum_cases <- purrr::map(by.location, function(df) {
    mRIIDS:::update_cases_column(df, case.type) 
})


outfile <- paste0(params$outfile, "_cumulative.csv")

bind_rows(cum_cases) %>%
    select(country, date, cases) %>%
readr::write_csv(
                 path = here::here("data/processed",
                                   outfile))

    
```

## Merge duplicate alerts

```{r data-cleanup-9 }
cols.to.keep <- c("location", "country", "disease", "species",
                  "healthmap_alert_id", "headline", "url",
                  "alert_tag", "feed_name", "lon", "lat")
## These are columns we generate ourselves later on
cols.to.keep <- c(cols.to.keep, "date", "cases") 
```

```{r data-cleanup-10 }
no_dups <- purrr::map(cum_cases, function(df) {
    mRIIDS:::merge_duplicates(df, cols.to.keep)
})    
```

## Remove outliers. 

```{r data-cleanup-11 }
use_last   <- 20
p.within.k <- 0.50
k_sd       <- mRIIDS:::interval_width_for_p(use_last,
                                            1 - p.within.k) %>%
    sqrt %>%
    `[`(2)

no_outliers <- purrr::map(no_dups, function(df) {
    df <- arrange(df, date)
    df <- filter(df, !is.na(cases)) %>% select(date, cases)
    if (nrow(df) == 0) {
        return(NULL)
    } else {
        df <- mRIIDS:::remove_last_outliers(df,
                                            use_last = use_last,
                                            k_sd = k_sd)
    }     
    message(df$country[1])
    df
})    

```

## Make monotonically increasing.
```{r data-cleanup-12 }

mono_inc <- purrr::map(no_outliers,
                       ~ mRIIDS:::make_monotonically_increasing(.x))

```

## Interpolate missing data
Finally, interpolate missing data.

```{r data-cleanup-13 }
interpolated <- purrr::map(mono_inc,
                           ~ mRIIDS:::interpolate_missing_data(.x,
                                                               method = "loglinear"))
    
  
```

## Plot


### Total cases as sum of sc and cc.

Common plotting logic. x must have a date and a cases column.

```{r data-cleanup-14}
cases_ts <- function(x) {
    p <- ggplot(x, aes(date, cases)) + geom_point()
    p <- p + theme_classic() 
    p <- p + xlab("") + ylab("Cases")
    p <- p + theme(axis.text.x = element_text(angle = 80, hjust = 1))
    p <- p + xlab("")
    p <- p + scale_x_date(date_labels =  "%b %Y")
    p     
}

```

```{r data-cleanup-15 }
total_p <- purrr::map(cum_cases, cases_ts)
```

### Remove duplicate alerts

```{r data-cleanup-16 }
nodups_p <- purrr::map(no_dups, cases_ts) 
```

## Remove outliers 

```{r data-cleanup-17 }
nooutss_p <- purrr::map(no_outliers, cases_ts)
```

### Make increasing

```{r data-cleanup-18 }
monoinc_p <- purrr::map(mono_inc, cases_ts)
```

### Interpolated data

This needs a slightly different viz.

```{r data-cleanup-19, eval = TRUE}
interp_p <- purrr::map(interpolated, function(x) {
    x$src <- "Raw"
    idx <- which(is.na(x$cases))
    x$src[idx] <- "Interpolated"
    x <- select(x,
                date = interpolated_date,
                cases = interpolated_cases,
                src)
    p <- ggplot(x, aes(date, cases, col = src)) + geom_point()
    p <- p + theme_classic() 
    p <- p + xlab("") + ylab("Cases")
    p <- p + theme(axis.text.x = element_text(angle = 80, hjust = 1))
    p <- p + theme(panel.border = element_blank(),
                   axis.line = element_line())
    p <- p + xlab("") + theme(legend.title = element_blank())
    p <- p + scale_x_date(date_labels =  "%b %Y")
    p     
})    
```

Putting them together.

```{r data-cleanup-20, eval = TRUE}

countries <- names(by.location)
graphs <- list(raw_p,
               total_p,
               nodups_p,
               nooutss_p,
               monoinc_p,
               interp_p)
purrr::map(countries, function(c){
    cspecific <- purrr::map(graphs, ~ .x[[c]])
    p <- ggpubr::ggarrange(plotlist = cspecific, nrow = 3, ncol = 2)
    ggpubr::ggexport(p,
                     filename = here::here("data/output/figures",
                                           paste0(params$outfile,
                                                 "_",
                                                 c,
                                                 ".png")),
                     res = 72,
                     width = 960,
                     height = 540)
})

```

## Derive incidence time series from interpolated cumulative cases

Stan is not going to like real numbers for a Poisson process.
So first we cast the cumulative case counts as integers and then 
take the difference.


```{r data-cleanup-21 }
incid_tall <- purrr::map_dfr(interpolated, function(x) {
    x <- select(x,
                date = interpolated_date,
                cases = interpolated_cases)
    x$cases <- as.integer(x$cases)
    x$incid <- c(0, diff(x$cases))
    x <- select(x, date, incid)
    x
}, .id = "country")                 
```

## Manually fix Ghana 

Manually fix the erroneous entry for Ghana on 2014-04-08


```{r ghana-fix}
incid_tall$country <- countrycode::countrycode(incid_tall$country,
                                               'country.name',
                                               'iso3c')

fix <- which(incid_tall$date == "2014-04-08" &
             incid_tall$country == "GHA")
incid_tall$incid[fix] <- 0
```

Also, get rid of the first row of all 0s to get Stan started.

```{r all-zeros}

incid_tall <- filter(incid_tall,
                     date != "2014-03-21")

```
And reshape.

```{r data-cleanup-22 }
incid_wide <- tidyr::spread(incid_tall, country, incid, fill = 0)
```


## Add 0 incidence for other countries

Finally, we add 0 incidence for all countries other than ones for
which we already have data. 

```{r data-cleanup-24}
ofinterest <- readr::read_csv(here::here(params$ofinterest)) %>%
    filter(include == "TRUE")

nodata <- setdiff(ofinterest$ISO3,
                  countrycode::countrycode(wafrica,
                                           'country.name',
                                           'iso3c'))
df <- matrix(0,
             nrow = nrow(incid_wide),
             ncol = length(nodata))
colnames(df) <- nodata
df <- data.frame(df)
incid_wide_all <- cbind(incid_wide, df)
```

Finally ensure that the order of countries in centroids and incid 
files is the same.

```{r data-cleanup-25, eval = TRUE}
inorder <- ofinterest$ISO3
## Some of the countries of interest may not be present in
## incid_wide_all
inorder <- intersect(inorder, colnames(incid_wide_all))
incid_wide_all <- select(incid_wide_all,
                         date,
                         inorder)
```


## Write the output

```{r data-cleanup-26 }

outfile <- paste0(params$outfile, "_wide.csv")
here::here("data/processed", outfile) %>%
    readr::write_csv(x = incid_wide_all,
                     path = .)

```
 
## Also make note of what points were interpolated.

```{r data-cleanup-27 }
incid_extracols <- purrr::map_dfr(interpolated, function(x) {
    idx <- which(is.na(x$cases))
    x$interpolated <- "FALSE"
    x$interpolated[idx] <- "TRUE"

    x <- select(x,
                date,
                interpolated)
},
.id = "country")

incid_extracols$country <- countrycode::countrycode(incid_extracols$country,
                                                    'country.name',
                                                    'iso3c')

```

Add the data that was affixed later.

```{r other-countries-tall}
df <- cbind(date = incid_wide_all$date, df)
df_tall <- tidyr::gather(df,
                         country,
                         incid,
                         -date)

incid_tall <- rbind(incid_tall,
                    df_tall)
```

```{r incidtall2}
incid_extracols <- left_join(incid_tall,
                             incid_extracols,
                             by = c("date",
                                    "country"))

outfile <- paste0(params$outfile, "_tall.csv")

here::here("data/processed", outfile) %>%
    readr::write_csv(x = incid_extracols,
                     path = .)

```

## Weekly series

Add a flag to indicate if all points in the week were reconstructed.

```{r data-cleanup-28 }
extra <- as.numeric(max(incid_extracols$date) -
                    min(incid_extracols$date)) %% 7

dates <- sort(incid_extracols$date) %>%
    unique

remove <- tail(dates, extra)

daily <- filter(incid_extracols,
                ! date %in% remove)
daily$interpolated <- as.logical(daily$interpolated)
weeks <- cut(daily$date, breaks = "7 days")
weekly <- split(daily, weeks) %>%
    purrr::map_dfr( ~ dplyr::group_by(.x,
                                      country) %>%
                        summarise(incid = sum(incid),
                                  interpolated = all(interpolated)),
                   .id = "date")

outfile <- paste0(params$outfile, "_weekly.csv")
readr::write_csv(x = weekly,
                 path = here::here("data/processed",
                                   outfile))





```
