---
output: 
pdf_document:
citation_package: natbib
keep_tex: false
fig_caption: true
latex_engine: pdflatex
title: Extract MCMC draws from the fitted model
author:
- name: Sangeeta Bhatia
affiliation: Imperial College London
abstract: 
keywords: 
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  tproj: 21
  twindow: 14
  N: 55
  day0: 2014-03-21
  indir: data/promed_stanfits_gamma_ul_10
---

```{r setup, eval = TRUE}
library(dplyr)
library(purrr)

```

## Read incidence data

The order of columns as the same as the order of columns in the input
to Stan. Hence read the incidence data to get this order.

```{r incid}
suffix <- paste0(params$tproj, 
                 "_", 
                 params$twindow, 
                 ".rds")

incid <- here::here(params$indir,
                    paste0("incid_",
                            suffix)) %>%
    readr::read_rds()

trng_date <- seq(from = as.Date(params$day0),
                 by = "1 day",
                 length.out = params$tproj)
  ##incid$date[seq_len(params$tproj)]
```

## Read in fit object

Extract the samples from the fitted Stan model and write them out for
later use. 

```{r fit}

fit1 <- readRDS(here::here(params$indir, suffix))
```

## Extract draws for each parameter

```{r draws}
list_of_draws <- rstan::extract(fit1)
r_samples     <- list_of_draws[["R"]]
gamma_samples <- list_of_draws[["gamma"]]
pstay_samples <- list_of_draws[["pstay"]]

```

## R index matrix 

For R, the matrix rindex contains the mapping of variable 
name to time window and country. Attach dates to the rindex data frame 
so we will know what variables 
are for which date.

```{r rindex}
rindex <- paste0("rindex_", suffix)
rindex <- here::here(params$indir,
                     rindex) %>% readRDS()

rindex_df <- as.data.frame(rindex)
rindex_df <- cbind(trng_date, rindex_df)
colnames(rindex_df) <- c("date", colnames(incid))
```

Each column in r_samples is a set of draws for a variable. The column
number identifies the variable number.

```{r}
rindex_df <- tidyr::gather(rindex_df, country, var, -date)
```

Now go over each row of this data frame, look at the var variable,
grab the corresponding column from r_samples and affix it to this row.

```{r reshape}
## samples <- purrr::pmap(rindex_df, ~ r_samples[, ..3])
samples <- purrr::pmap(rindex_df, function(date, country, var) r_samples[ , var])
samples <- do.call(rbind, samples)
rsamples <- cbind(rindex_df, samples)
```
Also calculate the quantiles for each row. Apply flips the rows and 
columns, hence some data wrangling is needed.

```{r}
sample_qntls <- apply(samples,
                      1,
                      quantile,
                      probs = c(0.025, 0.5, 0.975))
sample_qntls <- t(sample_qntls)
sample_qntls <- cbind(rindex_df, sample_qntls)
```

Finally save everything.

```{r write}
by_country <- split(rsamples, rsamples$country)
purrr::imap(by_country, ~ saveRDS(.x,
                           here::here(params$indir,
                                      paste0(.y,
                                             "_rsamples_",
                                             suffix))))


by_country <- split(sample_qntls, sample_qntls$country)
purrr::imap(by_country, ~ saveRDS(.x,
                           here::here(params$indir,
                                      paste0(.y,
                                             "_rquantiles_",
                                             suffix))))

saveRDS(pstay_samples,
        here::here(params$indir,
                   paste0("pstay_samples_", suffix)))


saveRDS(gamma_samples,
        here::here(params$indir,
                   paste0("gamma_samples_", suffix)))

        
```
