---
output: 
pdf_document:
citation_package: natbib
keep_tex: false
fig_caption: true
latex_engine: pdflatex
title: Extract MCMC draws from the fitted model
author:
- name: Sangeeta Bhatia
affiliation: Imperial College London
abstract: 
keywords: 
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  tproj: 112
  twindow: 28
  N: 3
  incid: healthmap_wide_from_july.csv
---

```{r setup, eval = TRUE}
library(dplyr)
library(purrr)

```

## Read incidence data

The order of columns as the same as the order of columns in the input
to Stan. Hence read the incidence data to get this order.

```{r incid}
incid <- here::here("data/processed", params$incid) %>%
    readr::read_csv(n_max = params$tproj)
trng_date <- incid$date[seq_len(params$tproj)]
```

## Read in fit object

Extract the samples from the fitted Stan model and write them out for
later use. 

```{r fit}
suffix <- paste0(params$tproj, "_", params$twindow, ".rds")
fit1 <- readRDS(here::here("data/stanfits", suffix))
```

## Extract draws for each parameter

```{r draws}
list_of_draws <- rstan::extract(fit1)
r_samples     <- list_of_draws[["R"]]
gamma_samples <- list_of_draws[["gamma"]]
pstay_samples <- list_of_draws[["pstay"]]

```

## R index matrix 

For R, the matrix rindex contains the mapping of variable 
name to time window and country.

```{r rindex}
rindex <- paste0("rindex_", suffix)
rindex <- here::here("data/stanfits",
                     rindex) %>% readRDS()

rindex_df <- as.data.frame(rindex)
colnames(rindex_df) <- colnames(select(incid, -date))
```

Attach dates to the rindex data frame so we will know what variables 
are for which date.


```{r}
rindex_df$date <- trng_date

```

Each column in r_samples is a set of draws for a variable. The column
number identifies the variable number.

```{r}
rindex_df <- tidyr::gather(rindex_df, country, var, -date)
```

Now go over each row of this data frame, look at the var variable,
grab the corresponding column from r_samples and affix it to this row.

```{r reshape}
samples <- purrr::pmap(rindex_df, ~ r_samples[, ..3])
samples <- do.call(rbind, samples)
rsamples <- cbind(rindex_df, samples)
```
Also calculate the quantiles for each row. Apply flips the rows and 
columns, hence some data wrangling is needed.

```{r}
sample_qntls <- apply(samples,
                      1,
                      quantile,
                      probs = c(0.025, 0.5, 0.975))
sample_qntls <- t(sample_qntls)
sample_qntls <- cbind(rindex_df, sample_qntls)
```

Finally save everything.

```{r write}
by_country <- split(rsamples, rsamples$country)
purrr::imap(by_country, ~ saveRDS(.x,
                           here::here("data/stanfits",
                                      paste0(.y,
                                             "_rsamples_",
                                             suffix))))


by_country <- split(sample_qntls, sample_qntls$country)
purrr::imap(by_country, ~ saveRDS(.x,
                           here::here("data/stanfits",
                                      paste0(.y,
                                             "_rquantiles_",
                                             suffix))))

saveRDS(pstay_samples,
        here::here("data/stanfits",
                   paste0("pstay_samples_", suffix)))


saveRDS(gamma_samples,
        here::here("data/stanfits",
                   paste0("gamma_samples_", suffix)))

        
```
