---
output: 
pdf_document:
citation_package: natbib
keep_tex: false
fig_caption: true
latex_engine: pdflatex
title: Visualisation of weekly performance metrics
author:
- name: Sangeeta Bhatia
affiliation: Imperial College London
abstract: 
keywords: 
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  twindow : 14
  n.dates.sim : 28
  places: !r c("SLE", "LBR", "GIN")
  indir: data/output
  outdir: data/output/
  datasource: ProMED
---

```{r setup, eval = TRUE}
library(dplyr)
library(purrr)
library(ggplot2)
source(here::here("analysis/common_plot_properties.R"))
```



```{r forecasts-assess-viz-1 }
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dev = "pdf",
  ## fig.width = 7,
  ## dev.args = list(pointsize = 6),
  fig.path = here::here(params$outdir),
  fig.ext = paste0(
    params$datasource,
    "_",
    params$twindow,
    "_",
    params$n.dates.sim,
    "_assess.pdf"
  )
)
```
Read in the files for all places of interest and visualise.

```{r forecasts-assess-viz-2 }

weekly_metrics <- readr::read_csv(
  file = here::here(
    params$indir,
    "weekly_metrics.csv"
  )
) %>% select(-logaccuracy)
```


## Proportion of observations in CI by week.

```{r}
infiles <- paste0(
  params$places,
  "_prop_in_ci_by_week.csv"
)

infiles <- here::here(
  paste0(
    params$indir,
    "/metrics/weekly"
  ),
  infiles
)
names(infiles) <- params$places

prop_in_ci_by_week <- purrr::map_dfr(
  infiles,
  ~ readr::read_csv(.x, col_names = FALSE),
  .id = "country"
)

colnames(prop_in_ci_by_week) <- c(
  "country",
  "tproj",
  "twindow",
  "n.dates.sim",
  "week_of_projection",
  "prop_in_ci"
)
```

Join it with R estimate categorisation.

```{r forecasts-assess-viz-3 }

epidemic_phase <- readr::read_csv(
  file = here::here(
    params$outdir,
    paste0(
      params$datasource,
      "_rquantiles_projection.csv"
    )
  )
)

epidemic_phase <- epidemic_phase[
  epidemic_phase$twindow == params$twindow,
]

epidemic_phase <- epidemic_phase[rep(
  seq_len(nrow(epidemic_phase)),
  each = params$n.dates.sim
), ]

## Fix the dates.
epidemic_phase <- dplyr::group_by(
  epidemic_phase,
  tproj,
  country
) %>%
  dplyr::mutate(
    date = seq(from = min(date), length.out = params$n.dates.sim, by = "1 day")
  )



weekly_metrics <- dplyr::left_join(
  weekly_metrics,
  epidemic_phase,
  by = c("tproj", "twindow", "country")
)

weekly_metrics <- weekly_metrics[
  weekly_metrics$twindow == params$twindow &
    weekly_metrics$n.dates.sim == params$n.dates.sim,
]
```

Overall performace

```{r}
weekly_metrics_overall <- weekly_metrics
weekly_metrics_overall$ci <- "Overall"
weekly_metrics <- rbind(weekly_metrics, weekly_metrics_overall)
```
## Fix column classes
```{r forecasts-assess-viz-4 }

weekly_metrics <- dplyr::mutate_if(
  weekly_metrics,
  is.character,
  factor
)
weekly_metrics$week_of_projection <- factor(weekly_metrics$week_of_projection)

weekly_metrics$ci <- forcats::fct_recode(
  weekly_metrics$ci,
  "Growing" = "ll_greater_than_1",
  "Declining" = "ul_less_than_1",
  "Neither" = "ci_includes_1",
  "Overall" = "Overall"
  )

weekly_metrics$ci <- forcats::fct_relevel(
  weekly_metrics$ci,
  c("Growing",
    "Declining",
    "Neither",
    "Overall"
   )   
)
                                  
weekly_metrics$tproj <- as.integer(weekly_metrics$tproj)
weekly_metrics$twindow <- factor(weekly_metrics$twindow)
```


Same for prop in ci.

```{r}
prop_in_ci_by_week <- dplyr::left_join(
  prop_in_ci_by_week,
  epidemic_phase,
  by = c("tproj", "twindow", "country")
)

prop_in_ci_by_week <- prop_in_ci_by_week[
  prop_in_ci_by_week$twindow == params$twindow &
    prop_in_ci_by_week$n.dates.sim == params$n.dates.sim,
]
```

Overall performance

```{r}
prop_in_ci_by_week_overall <- prop_in_ci_by_week
prop_in_ci_by_week_overall$ci <- "Overall"
prop_in_ci_by_week <- rbind(
    prop_in_ci_by_week,
    prop_in_ci_by_week_overall
)

prop_in_ci_by_week$ci <- forcats::fct_recode(
  prop_in_ci_by_week$ci,
  "Growing" = "ll_greater_than_1",
  "Declining" = "ul_less_than_1",
  "Neither" = "ci_includes_1",
  "Overall" = "Overall"
  )

prop_in_ci_by_week$ci <- forcats::fct_relevel(
  prop_in_ci_by_week$ci,
  c("Growing",
    "Declining",
    "Neither",
    "Overall"
   )   
)

```


```{r forecasts-assess-viz-5 }

summary_stats <- summarise_at(
  weekly_metrics,
  vars(rmae:bias),
  list(
    avg = ~ mean(., na.rm = TRUE),
    med = ~ quantile(., 0.5, na.rm = TRUE),
    low = ~ quantile(., 0.25, na.rm = TRUE),
    high = ~ quantile(., 0.75, na.rm = TRUE)
  )
)

summary_stats <- tidyr::gather(
  summary_stats,
  key = metric,
  value = value
) %>%
  tidyr::separate(
    col = metric,
    into = c("metric", "bound")
  ) %>%
  tidyr::spread(
    key = bound,
    value = value,
    convert = TRUE
  )

select(
  summary_stats,
  metric,
  avg,
  med,
  low,
  high
) %>% knitr::kable()
```

## Summary stats by week


```{r forecasts-assess-viz-6 }

weekly_summary_stats <- group_by(weekly_metrics, week_of_projection) %>%
  summarise_at(
    vars(rmae:bias),
    list(
      avg = ~ mean(., na.rm = TRUE),
      med = ~ quantile(., 0.5, na.rm = TRUE),
      low = ~ quantile(., 0.25, na.rm = TRUE),
      high = ~ quantile(., 0.75, na.rm = TRUE)
    )
  )
```

```{r forecasts-assess-viz-7}
knitr::kable(weekly_summary_stats)
```

## Summary stats by week and phase of epidemic

```{r forecasts-assess-viz-8}

phase_weekly_summary_stats <- group_by(
  weekly_metrics,
  week_of_projection,
  ci
) %>%
  summarise_at(
    vars(rmae:bias),
    list(
      avg = ~ mean(., na.rm = TRUE),
      med = ~ quantile(., 0.5, na.rm = TRUE),
      low = ~ quantile(., 0.25, na.rm = TRUE),
      high = ~ quantile(., 0.75, na.rm = TRUE)
    )
  )

phase_weekly_summary_stats <- tidyr::gather(
  phase_weekly_summary_stats,
  key = metric,
  value = value,
  -week_of_projection,
  -ci
) %>%
  tidyr::separate(
    col = metric,
    into = c("metric", "bound")
  ) %>%
  tidyr::spread(
    key = bound,
    value = value,
    convert = TRUE
  )

phase_weekly_summary_stats <- dplyr::mutate_if(
  phase_weekly_summary_stats,
  is.numeric,
  round,
  3
)

phase_weekly_summary_stats$formatted_iqr <- paste0(
  phase_weekly_summary_stats$med,
  " (",
  phase_weekly_summary_stats$low,
  " - ",
  phase_weekly_summary_stats$high,
  ")"
)

phase_weekly_summary_stats <- dplyr::select(
  phase_weekly_summary_stats,
  week_of_projection,
  ci,
  metric,
  formatted_iqr
) %>%
  tidyr::spread(
    key = metric,
    value = formatted_iqr
  )
```

```{r forecasts-assess-viz-9}
knitr::kable(phase_weekly_summary_stats)
```
## Bias

```{r forecasts-assess-viz-10 }
pbias <- ggplot(
  weekly_metrics,
  aes(week_of_projection,
    bias,
    col = ci
  )
) +
  geom_boxplot(position = "dodge")
pbias <- pbias +
  xlab("Week of projection") +
  ylab("Bias") +
  mriids_plot_theme$theme +
  mriids_plot_theme$legend +
  scale_color_manual(values = mriids_plot_theme$phase_color_scale)

pbias
```

## Sharpness


```{r forecasts-assess-viz-11 }

psharp <- ggplot(
  weekly_metrics,
  aes(week_of_projection,
    sharpness,
    col = ci
  )
) +
    geom_boxplot() +
    scale_y_log10(
         breaks = scales::trans_breaks("log10", function(x) 10^x),
         labels = scales::trans_format("log10", scales::math_format(10^.x)))


psharp <- psharp +
  xlab("Week of projection") +
  ylab("log(sharpness)") +
  mriids_plot_theme$theme +
  mriids_plot_theme$legend +
  scale_color_manual(values = mriids_plot_theme$phase_color_scale)

psharp
```

## Proportion of observations in CrI


```{r forecasts-assess-viz-12 }
prop_in_ci_by_week$week_of_projection <- factor(prop_in_ci_by_week$week_of_projection)
prel <- ggplot(
  prop_in_ci_by_week,
  aes(week_of_projection,
    prop_in_ci,
    col = ci
  )
) +
  geom_boxplot()
## geom_point(aes(fill = ci),
##            alpha = 0.7,
##            position = position_jitterdodge())
prel <- prel +
  xlab("Week of projection") +
  ylab("Proportion in 95% forecasts interval") +
  mriids_plot_theme$theme +
  mriids_plot_theme$legend +
  scale_color_manual(values = mriids_plot_theme$phase_color_scale)

prel
```


## Relative mean absolute error


```{r forecasts-assess-viz-13 }
prelerr <- ggplot(
  weekly_metrics,
  aes(week_of_projection,
    rmae,
    col = ci
  )
) +
    geom_boxplot() +
    scale_y_log10(
         breaks = scales::trans_breaks("log10", function(x) 10^x),
         labels = scales::trans_format("log10", scales::math_format(10^.x)))

prelerr <- prelerr +
  mriids_plot_theme$theme +
  mriids_plot_theme$legend +
  scale_color_manual(values = mriids_plot_theme$phase_color_scale)


prelerr <- prelerr +
  xlab("Week of projection") +
  ylab("log(relative mean absolute error)")

prelerr
```


```{r forecasts-assess-viz-14 }
outfile <- here::here(
  params$outdir,
  paste0(
    "forecasts_assess_",
    params$datasource,
    params$twindow,
    ".png"
  )
)

## Affix legend manually otherwise I can't get rid of the
## annoying legend title.
legend <- cowplot::get_legend(pbias +
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  ))

prow <- cowplot::plot_grid(pbias + theme(legend.position = "none"),
  psharp + theme(legend.position = "none"),
  prel + theme(legend.position = "none"),
  prelerr + theme(legend.position = "none"),
  align = "vh",
  labels = c("A", "B", "C", "D"),
  label_size = 8,
  hjust = c(-1, 1, -3, 1),
  vjust = c(3, 3, -2.5, 0),
  nrow = 2
)


# add the legend underneath the row we made earlier. Give it 10% of the height
# of one plot (via rel_heights).
p <- cowplot::plot_grid(prow,
  legend,
  ncol = 1,
  rel_heights = c(1, .1)
)

p
```
