---
output: 
pdf_document:
citation_package: natbib
keep_tex: false
fig_caption: true
latex_engine: pdflatex
title: 
author:
- name: Sangeeta Bhatia
affiliation: Imperial College London
abstract: 
keywords: 
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  tproj : 49
  twindow : 42
  n.dates.sim : 28
  incid: data/processed/promed_loglinear_wide_from_march.csv
  place: LBR
---

```{r setup, eval = TRUE}
library(dplyr)
library(purrr)
library(ggplot2)
```
The metrics are implemented in R package assessR.

```{r install, eval = FALSE}
devtools::install_github("sangeetabhatia03/assessR")
```

Read in the forecasts and the observed incidence.

```{r incid}
incid <- here::here(params$incid) %>%
    readr::read_csv()

idx <- seq(from = params$tproj + 1,
           length.out = params$n.dates.sim,
           by = 1)

observed <- incid[idx, ]

```

```{r pred}
infile <- paste0("forecasts_samples_",
                 params$tproj,
                  "_",
                 params$twindow,
                 "_",
                 params$n.dates.sim,
                 ".Rds")
predicted <- here::here("data/output",
                        infile) %>%
    readr::read_rds()

```
We will need to reshape the forecasts so that we can use the functions
in assessR.

```{r reshape}
predicted <- bind_cols(predicted)
pred <- select(predicted, starts_with(params$place)) %>%
    as.matrix
obs <- pull(observed, params$place)
```

Calculate performance along the four metrics we have selected.

```{r}
rmae <- assessR::rel_mae(obs, pred)
## avg_res <- assessR::avg_residual(obs, pred)
sness <- assessR::sharpness(pred)
bs <- assessR::bias(obs, pred)

```

And the proportion of observed values in 50% CI.

```{r propinci}
minmax <- apply(pred, 1, quantile, c(0.25, 0.75))
minmax <- t(minmax)
propinci <- assessR:::prop_in_ci(obs,
                                 min = minmax[, 1],
                                 max = minmax[, 2])
df <- data.frame(tproj = params$tproj,
                 twindow = params$twindow,
                 propinci = propinci
                 )
outfile <- paste0(params$place, 
                 "_prop_in_ci.csv")

here::here("data/output/metrics/daily",
           outfile) %>%
    readr::write_csv(x = df,
                     path = .,
                     append = TRUE)

```

Finally,  save the output for later analysis and visualisation.

```{r save}
df <- data.frame(date = observed$date,
                 rmae = rmae,
                 sharpness = sness,
                 bias = bs)

outfile <- paste0(params$place, 
                  "_daily_metrics_",
                  params$tproj,
                  "_",
                  params$twindow,
                  "_",
                  params$n.dates.sim,
                  ".csv")
here::here("data/output/metrics/daily",
           outfile) %>%
    readr::write_csv(x = df,
                     path = .)

```

Also save the averages while we are at it.

```{r}
weekly <- mRIIDS:::daily.to.weekly(df)
df_weekly <- mutate_if(weekly, is.numeric, funs(. / 7))

outfile <- paste0(params$place, 
                  "_weekly_metrics_",
                  params$tproj,
                  "_",
                  params$twindow,
                  "_",
                  params$n.dates.sim,
                  ".csv")
here::here("data/output/metrics/weekly",
           outfile) %>%
    readr::write_csv(x = df_weekly,
                     path = .)

```

```{r}
devtools::session_info()
```
