---
output: 
pdf_document:
citation_package: natbib
keep_tex: false
fig_caption: true
latex_engine: pdflatex
title: Visualisation of weekly performance metrics
author:
- name: Sangeeta Bhatia
affiliation: Imperial College London
abstract: 
keywords: 
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  twindow : 14
  n.dates.sim : 28
  places: !r c("SLE", "LBR", "GIN")
  indir: data/output
  outdir: data/output/
  datasource: ProMED
---

```{r setup, eval = TRUE}
library(dplyr)
library(purrr)
library(ggplot2)
source(here::here("analysis/common_plot_properties.R"))
```



```{r forecasts-assess-viz-1 }
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dev = "pdf",
  fig.path = here::here(glue::glue("{params$outdir}/{Sys.Date()}/")),
  fig.width = mriids_plot_theme$single_col_width / 2.5,
  fig.height = mriids_plot_theme$single_col_height / 2.5,
  fig.ext = paste0(
    params$datasource,
    "_",
    params$twindow,
    "_",
    params$n.dates.sim,
    "_assess.pdf"
  )
)
 ```
Read in the files for all places of interest and visualise.

```{r forecasts-assess-viz-2 }

weekly_metrics <- readr::read_csv(
  file = here::here(
    params$indir,
    glue::glue("{params$datasource}_weekly_metrics.csv")
  )
)

weekly_metrics <- weekly_metrics[
  weekly_metrics$twindow == params$twindow &
    weekly_metrics$n.dates.sim == params$n.dates.sim,
  ]

weekly_metrics$twindow <- as.factor(weekly_metrics$twindow)
## to avoid problems when we split on underscore.
weekly_metrics <-  dplyr::rename(weekly_metrics, relsharpness = "rel_sness")
```

## Overall performance

```{r forecasts-assess-viz-3 }

skimr::skim_with(
           numeric = list(
               low = ~quantile(., probs = 0.025),
               high = ~quantile(., probs = 0.975),
               missing = NULL,
               complete = NULL
           ),    
          append = TRUE
       )

df <- skimr::skim(weekly_metrics, rmae:bias)

out <- glue::glue(
  "{params$datasource}_weekly_metrics_summary_",
  "{params$twindow}_{params$n.dates.sim}.csv"
)
readr::write_csv(x = df, path = here::here(params$outdir, out))
```

```{r forecasts-assess-viz-4 }
skimr::kable(df)
```


Join weekly metrics with R estimate categorisation.

```{r forecasts-assess-viz-5 }

epidemic_phase <- readr::read_csv(
  file = here::here(
    params$outdir,
    paste0(
      params$datasource,
      "_rquantiles_projection.csv"
    )
  )
)

epidemic_phase <- epidemic_phase[
  epidemic_phase$twindow == params$twindow,
]

## epidemic_phase <- epidemic_phase[rep(
##   seq_len(nrow(epidemic_phase)),
##   each = params$n.dates.sim / 7
## ), ]

## Fix the dates.
## epidemic_phase <- dplyr::group_by(
##   epidemic_phase,
##   tproj,
##   country
## ) %>%
##   dplyr::mutate(
##     date = seq(from = min(date), length.out = params$n.dates.sim / 7, by = "7 days")
##    )

epidemic_phase$twindow <- as.factor(epidemic_phase$twindow)

weekly_metrics <- dplyr::left_join(
  weekly_metrics,
  epidemic_phase,
  by = c("tproj", "twindow", "country")
)
```

Check for NAs

```{r forecasts-assess-viz-6}
naci <- dplyr::filter(weekly_metrics, is.na(ci))
out <- glue::glue(
  "{params$datasource}_NA_ci_",
  "{params$twindow}_{params$n.dates.sim}.csv"
)
readr::write_csv(x = naci, path = here::here(params$outdir, out))
weekly_metrics <- dplyr::filter(weekly_metrics, !is.na(ci))
```

Overall performace

```{r forecasts-assess-viz-7 }
weekly_metrics_overall <- weekly_metrics
weekly_metrics_overall$ci <- "Overall"
weekly_metrics <- rbind(weekly_metrics, weekly_metrics_overall)
```

## Fix column classes
```{r forecasts-assess-viz-8 }

weekly_metrics <- dplyr::mutate_if(
  weekly_metrics,
  is.character,
  factor
)
weekly_metrics$week_of_projection <- factor(
    weekly_metrics$week_of_projection
)

weekly_metrics$ci <- forcats::fct_recode(
  weekly_metrics$ci,
  "Growing" = "ll_greater_than_1",
  "Declining" = "ul_less_than_1",
  "Stable" = "ci_includes_1",
  "Overall" = "Overall"
)

weekly_metrics$ci <- forcats::fct_relevel(
  weekly_metrics$ci,
  c(
    "Growing",
    "Declining",
    "Stable",
    "Overall"
  )
)

weekly_metrics$tproj <- as.integer(weekly_metrics$tproj)
weekly_metrics$twindow <- factor(weekly_metrics$twindow)
```

## Summary

```{r forecasts-assess-viz-9 }

summary_stats <- dplyr::summarise_at(
  weekly_metrics,
  vars(rmae:bias),
  list(
    avg = ~ mean(., na.rm = TRUE),
    med = ~ quantile(., 0.5, na.rm = TRUE),
    low = ~ quantile(., 0.25, na.rm = TRUE),
    high = ~ quantile(., 0.75, na.rm = TRUE)
  )
)

summary_stats <- tidyr::gather(
  summary_stats,
  key = metric,
  value = value
) %>%
  tidyr::separate(
    col = metric,
    into = c("metric", "bound")
  ) %>%
  tidyr::spread(
    key = bound,
    value = value,
    convert = TRUE
  )

dplyr::select(
  summary_stats,
  metric,
  avg,
  med,
  low,
  high
) %>% knitr::kable()
```

## Summary stats by week


```{r forecasts-assess-viz-10 }

weekly_summary_stats <- dplyr::group_by(weekly_metrics, week_of_projection) %>%
  dplyr::summarise_at(
    vars(rmae:bias),
    list(
      avg = ~ mean(., na.rm = TRUE),
      med = ~ quantile(., 0.5, na.rm = TRUE),
      low = ~ quantile(., 0.25, na.rm = TRUE),
      high = ~ quantile(., 0.75, na.rm = TRUE)
    )
  )
```

```{r forecasts-assess-viz-11 }
knitr::kable(weekly_summary_stats)
```


## Summary stats by week and phase of epidemic

```{r forecasts-assess-viz-12 }

phase_weekly_summary_stats <- dplyr::group_by(
  weekly_metrics,
  week_of_projection,
  ci
) %>%
  dplyr::summarise_at(
    vars(rmae:bias),
    list(
      avg = ~ mean(., na.rm = TRUE),
      med = ~ quantile(., 0.5, na.rm = TRUE),
      low = ~ quantile(., 0.25, na.rm = TRUE),
      high = ~ quantile(., 0.75, na.rm = TRUE)
    )
  )

phase_weekly_summary_stats <- tidyr::gather(
  phase_weekly_summary_stats,
  key = metric,
  value = value,
  -week_of_projection,
  -ci
) %>%
  tidyr::separate(
    col = metric,
    into = c("metric", "bound")
  ) %>%
  tidyr::spread(
    key = bound,
    value = value,
    convert = TRUE
  )

phase_weekly_summary_stats <- dplyr::mutate_if(
  phase_weekly_summary_stats,
  is.numeric,
  round,
  3
)

phase_weekly_summary_stats$formatted_iqr <- paste0(
  phase_weekly_summary_stats$med,
  " (",
  phase_weekly_summary_stats$low,
  " - ",
  phase_weekly_summary_stats$high,
  ")"
)

phase_weekly_summary_stats <- dplyr::select(
  phase_weekly_summary_stats,
  week_of_projection,
  ci,
  metric,
  formatted_iqr
) %>%
  tidyr::spread(
    key = metric,
    value = formatted_iqr
  )
```

```{r forecasts-assess-viz-13 }
knitr::kable(phase_weekly_summary_stats)
```

## Proportion of observations in forecast interval

```{r forecasts-assess-viz-14}
propinci <- group_by(weekly_metrics, country, ci) %>%
    summarise(x = mean(propinci))

propinci <- tidyr::spread(propinci, ci, x)

```

```{r forecasts-assess-viz-15}
knitr::kable(propinci)
```
and across the three countries.

```{r forecasts-assess-viz-16}
propinci_all <- group_by(weekly_metrics, ci) %>%
    summarise(x = mean(propinci))
propinci_all <- tidyr::spread(propinci_all, ci, x)
```


```{r forecasts-assess-viz-17}
knitr::kable(propinci_all)
```
Format and save.

```{r forecasts-assess-viz-18}
propinci_all$Country <- "All"
propinci_all <- dplyr::select(propinci_all, Country, dplyr::everything())
propinci$country <- countrycode::countrycode(propinci$country, "iso3c", "country.name")
propinci <- dplyr::ungroup(propinci)

propinci <- dplyr::rename(propinci, "Country" = "country")
propinci <- rbind(propinci, propinci_all)
propinci <- dplyr::mutate_if(propinci, is.numeric, scales::percent)

readr::write_csv(
    x = propinci,
    path = here::here(
        glue::glue("rmd/{params$datasource}_propinci_{params$twindow}",
             "_{params$n.dates.sim}.csv"
         )
    )
)

```
## Bias

```{r forecasts-assess-viz-19 }
pbias <- ggplot(
  weekly_metrics,
  aes(week_of_projection,
    bias,
    col = ci
  )
) +
  geom_boxplot(outlier.size = 1, outlier.stroke = 0.1, position = "dodge")
pbias <- pbias +
  xlab("") +
  ylab("bias") +
  mriids_plot_theme$theme +
  mriids_plot_theme$legend +
  theme(
        axis.title = ggplot2::element_text(margin = margin(t = 0, r = 0, b = 0, l = 0) , size = 6)
    ) +
  scale_color_manual(values = mriids_plot_theme$phase_color_scale)


pbias
```

## Relative Sharpness


```{r forecasts-assess-viz-20 }

psharp <- ggplot(
  weekly_metrics,
  aes(week_of_projection,
    relsharpness,
    col = ci
  )
) +
  geom_boxplot(outlier.size = 1, outlier.stroke = 0.1) 
  ## scale_y_log10(
  ##   breaks = scales::trans_breaks("log10", function(x) 10^x),
  ##   labels = scales::trans_format("log10", scales::math_format(10^.x))
  ## )


psharp <- psharp +
  xlab("") +
  ylab("sharpness") +
  mriids_plot_theme$theme +
  mriids_plot_theme$legend +
  theme(
        axis.title = ggplot2::element_text(margin = margin(t = 0, r = 0, b = 0, l = 0) , size = 6)
    ) +
  scale_color_manual(values = mriids_plot_theme$phase_color_scale)

psharp
```

## Proportion of observations in CrI


```{r forecasts-assess-viz-21 }
propinci_perc <- group_by(
    weekly_metrics,
    week_of_projection,
    ci
) %>% summarise(nweeks = 100 * sum(propinci) / n())
propinci_perc <- dplyr::ungroup(propinci_perc)
prel <- ggplot(
  propinci_perc,
  aes(week_of_projection,
    nweeks,
    fill = ci
  )
) +
  geom_bar(stat = "identity", position = "dodge", width = 0.5)
## geom_point(aes(fill = ci),
##            alpha = 0.7,
##            position = position_jitterdodge())
prel <- prel +
  xlab("") +
  ylab(stringr::str_wrap("% weeks in 95% forecast interval", width = 15)) +
  mriids_plot_theme$theme +
  mriids_plot_theme$legend +
  theme(
        axis.title = ggplot2::element_text(margin = margin(t = 0, r = 0, b = 0, l = 0) , size = 6)
    ) +
  ylim(0, 100) +
  scale_fill_manual(values = mriids_plot_theme$phase_color_scale)

prel
```


## Relative mean absolute error


```{r forecasts-assess-viz-22 }

prelerr <- ggplot(
  weekly_metrics,
  aes(week_of_projection,
    rmae,
    col = ci
  )
) +
  geom_boxplot(outlier.size = 1, outlier.stroke = 0.1) 

prelerr <- prelerr +
  mriids_plot_theme$theme +
  mriids_plot_theme$legend +
  scale_color_manual(values = mriids_plot_theme$phase_color_scale) +
  theme(
        axis.title = ggplot2::element_text(margin = margin(t = 0, r = 0, b = 0, l = 0) , size = 6)
    )


prelerr <- prelerr +
  xlab("") +
  ylab("relative mean absolute error")

prelerr
```

```{r forecasts-assess-viz-23}
prelerr_log <- prelerr +
    ##scale_y_continuous(trans = "log10") +
    ##annotation_logticks(sides = "l") +
      scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(.x))) +
     ylab("log(rmae)") +   theme(
        axis.title = ggplot2::element_text(margin = margin(t = 0, r = 0, b = 0, l = 0), size = 6)
    ) 

```

```{r forecasts-assess-viz-24 }
## outfile <- here::here(
##   params$outdir,
##   paste0(
##     "forecasts_assess_",
##     params$datasource,
##     params$twindow,
##     ".pdf"
##   )
## )

## Affix legend manually otherwise I can't get rid of the
## annoying legend title.
legend <- cowplot::get_legend(pbias +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 6)
  ))

prow <- cowplot::plot_grid(
  prel + theme(legend.position = "none"),
  prelerr + theme(legend.position = "none"),
  pbias + theme(legend.position = "none"),
  psharp + theme(legend.position = "none"),
  align = "vh",
  labels = c("A", "B", "C", "D"),
  label_size = 8,
  hjust = c(-1, 1, -3, 1),
  vjust = c(3, 3, -2.5, 0),
  nrow = 2
)


# add the legend underneath the row we made earlier. Give it 10% of the height
# of one plot (via rel_heights).
p <- cowplot::plot_grid(prow,
  legend,
  ncol = 1,
  rel_heights = c(1, .1)
)

p
```

And with log scale.

```{r forecasts-assess-viz-25}
prow <- cowplot::plot_grid(
  prel + theme(legend.position = "none"),
  prelerr_log + theme(legend.position = "none"),
  pbias + theme(legend.position = "none"),
  psharp + theme(legend.position = "none"),
  align = "vh",
  labels = c("A", "B", "C", "D"),
  label_size = 8,
  hjust = c(-1, 1, -3, 1),
  vjust = c(3, 3, -2.5, 0),
  nrow = 2
)


# add the legend underneath the row we made earlier. Give it 10% of the height
# of one plot (via rel_heights).
p <- cowplot::plot_grid(prow,
  legend,
  ncol = 1,
  rel_heights = c(1, .1)
)

p
```
