---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
title: Estimation of Model Parameters
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "`r Sys.Date()`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  tproj : 329
  twindow : 7
  n.dates.sim : 35
  mobilty: !r list(K = 1, pow_N_from = 1, pow_N_to = 1)
  incid: 
  nsamples: 1000  
---


```{r setup, eval = TRUE}

library(ggthemes)
library(ggplot2)
library(dplyr)
library(rstan)
library(EpiEstim)
library(mRIIDS)
```

## Read incidence data

```{r incid}
incid <- here::here("data/processed", params$incid) %>%
    readr::read_csv()
trng <- incid[seq_len(params$tproj), ]
valdtn <- incid[(params$tproj + 1):nrow(incid), ]

I0    <- as.matrix(select(trng, -date))
n_loc <- ncol(I0)

```

## Read stan fit object

```{r}
fitobj <- paste0(params$tproj, "_", params$twindow, ".rds")
fit1 <- here::here("data/stanfits", fitobj) %>%
    readRDS(.)
```

## R index matrix 



```{r rindex}
rindex <- paste0("rindex_",
                 params$tproj,
                 "_",
                 params$twindow,
                 ".rds")
rindex <- here::here("data/stanfits",
                     rindex) %>% readRDS()
```

## Ebola params
```{r}
si <- paste0("si_",
             params$tproj,
             "_",
             params$twindow,
             ".rds")
si <- here::here("data/stanfits",
                 si) %>%
    readRDS()
```
## Test model fit

The fitted model has draws from the posterior distribution.
We can use these samples to project forward thus doing either
goodness-of-fit test or predictive modeling.

```{r test_fit, eval = TRUE}

list_of_draws <- rstan::extract(fit1)
r_samples     <- list_of_draws[["R"]]
gamma_samples <- list_of_draws[["gamma"]]
pstay_samples <- list_of_draws[["pstay"]]

```

Distances have been calculated using geosphere package. Read in the 
distance matrix and extract the lower triangle.

```{r}

distances <- here::here("data/processed", "distances.mat") %>%
    read.table(header = FALSE)
distances <- as.matrix(distances) 
distances <- distances[lower.tri(distances)]

```

Read in populations and arrange then in the same order as in the
distance vector.

```{r}
centroids <- here::here("data/processed",
                        "centroids.csv") %>%
    read.csv()

idx <- mRIIDS:::lower_tri_idx(nrow(distances))
n_from <- centroids$pop[idx[, 1]]
n_to <- centroids$pop[idx[, 2]]

```

```{r project, eval = TRUE}
index <- sample(seq_len(nrow(r_samples)),
                params$nsamples)

mobilty <- params$mobilty
daily_projections <- lapply(index, function(row){
                                     gamma <- gamma_samples[row]
                                     pstay <- pstay_samples[row]
                                     R_posterior <- apply(rindex, c(1, 2),
                                                          function(i) r_samples[row, i])
                                     mobilty$pow_dist <- gamma
                                     flowmat <- mRIIDS:::flow_matrix(distances,
                                                            n_from,
                                                            n_to,
                                                            place_names = centroids$country,
                                                            model = "gravity",
                                                            mobilty)


                                     ## Relative risk
                                     rel_risk <- flowmat / rowSums(flowmat, na.rm=TRUE)

                                     ## matrix characterising the population movement between geographical units

                                     pij <- mRIIDS:::probability_movement(rel_risk,
                                                                          pstay)
                                     
                                     out <- mRIIDS:::project(I0,
                                                             R_posterior,
                                                             SI_Distr,
                                                             pij,
                                                             n_days) 
                                     out <- 
                                      as.data.frame(out) 
                                     
                                     out$date <- tail(dates, nrow(out))
                                     out
})


```

```{r weekly, eval = !simulated_data}

weekly_projections <- lapply(daily_projections, daily.to.weekly) %>%
                      bind_rows(.)
projections_distr <- projection_quantiles(weekly_projections)

```

```{r}
weekly.available <- c(training    =
                          list(who_wide[seq_len(t.proj), ]),
                      validation =
                          list(who_wide[t.proj:(t.proj + n_days), ])) %>%
                       lapply(daily.to.weekly) %>%
                       bind_rows(.id = "Category")

trng.start <- pull(who_wide, Date)[t.proj] - time_window
vldtn.end  <- pull(who_wide, Date)[t.proj] + n_days
p <- plot.weekly3(weekly.available,
                  projections_distr, trng.start,
                  vldtn.end)

```

```{r plot_save, eval = TRUE}

paste0(ADM0, "-",
       t.proj, "-",
       n_days, "-", time_window, ".csv") %>%
    here::here("output", .) %>%
    readr::write_csv(x = projections_distr, path = .)

outname <- paste0(ADM0, "-", t.proj, "-", n_days, "-", time_window, ".png")
outfile <- here::here("output/figures", outname)
ggsave(outfile, p)


```


```{r test_sim, eval = simulated_data}

I_df <- as.data.frame(I)
I_df %<>% tibble::rownames_to_column(.)
I_tall <- tidyr::gather(I_df, var, val, -rowname)
I_tall$rowname %<>% as.numeric
ggplot(I_tall, aes(rowname, val)) + geom_point() +
    facet_grid(var~.) +
    geom_vline(xintercept = change_at)

```


