---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
title: Estimation of Model Parameters
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "`r Sys.Date()`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  tproj : 63
  twindow : 28
  n.dates.sim : 28
  day0: 2014-03-22
  nsamples: 900
  N: 55
  indir: data/who_stanfits
  outdir: data/who_output
---


```{r setup, eval = TRUE}

library(dplyr)
library(rstan)
library(mRIIDS)
```

## Read incidence data

```{r incid}
suffix <- paste0(params$tproj, 
                 "_", 
                 params$twindow, 
                 ".rds")

trng <- here::here(params$indir,
                    paste0("incid_",
                            suffix)) %>%
    readr::read_rds()

trng_dates <- seq(from = as.Date(params$day0),
                  by = "1 day",
                  length.out = params$tproj)

pred_dates <- seq(from = max(trng_dates) + 1,
                  length.out = params$n.dates.sim,
                  by = "1 day")
I0    <- as.matrix(trng)
n_loc <- ncol(I0)
countries <- colnames(trng)
```



## Ebola params
```{r ebpars}
suffix <- paste0(params$tproj, "_", params$twindow, ".rds")
si <- paste0("si_", suffix)
si <- here::here(params$indir,
                 si) %>%
    readr::read_rds()
```
## Samples from posterior distribution

The fitted model has draws from the posterior distribution.
We can use these samples to project forward thus doing either
goodness-of-fit test or predictive modeling.

```{r test_fit, eval = TRUE}
r_samples <- purrr::map(countries,
                       ~ readRDS(here::here(params$indir,
                                            paste0(.x,
                                                   "_rsamples_",
                                                   suffix))))

```

We want to use the estimate of R in the window right before we start
projecting as the R for the window over which we project. So we will
just pick out the relevant row from the samples.

```{r rpred}
last_window <- tail(trng_dates, 1)
r_pred <- purrr::map_dfr(r_samples, ~ filter(.x,
                                             date == last_window))
r_pred <- select(r_pred, - (date:var))
```

Now we will sample from the full draw for each parameter of the model.

```{r project, eval = TRUE}
nrsamples <- ncol(r_samples[[1]]) - 3 ## date, country, var in the way we have organised this.
index <- sample(seq_len(nrsamples),
                params$nsamples)
```

Flow matrices have been calculated on the cluster. However, they have been named 
sequentially from 1 to 900. So we can't use a random integer to call them.
To get around this problem, we will name each item using the random integers drawn in the previous step
and use this name to use the matrix.

```{r}
flow_mats <- here::here(params$indir,
                         paste0("flow_matrices/flow_",
                                seq_len(params$nsamples),
                                "_",
                                suffix)) %>%
  purrr::map(~ readr::read_rds(.x))
names(flow_mats) <- index
```

Now we are ready to project.

```{r}
daily_projections <- lapply(index, function(row){
                                     R_posterior <- r_pred[, row]
                                     R_posterior <- matrix(R_posterior,
                                                           ncol = ncol(I0),
                                                           nrow = 1)
                                     pij <- flow_mats[[as.character(row)]]
                                     out <- mRIIDS:::project(I0,
                                                             R_posterior,
                                                             si,
                                                             pij,
                                                             params$n.dates.sim) 
                                     out <- 
                                      as.data.frame(out) 
                                     
                                     out$date <- pred_dates
                                     out
})


```

```{r weekly, eval = TRUE}

weekly_projections <- purrr::map_dfr(daily_projections,
                                     mRIIDS:::daily.to.weekly) 
projections_distr <- mRIIDS:::projection_quantiles(weekly_projections)

```

## Aggregate projections for West African Region

```{r agg, eval = FALSE}

countries <- select(incid, -date) %>% colnames
wafrica_all <- purrr::map_dfr(countries, function(x) {
    out <- mRIIDS:::sample_projections(daily_projections,
                                       x,
                                       params$nsamples)
  out$date <- pred_dates
  out
})

aggregated <- split(wafrica_all, wafrica_all$date) %>%
  purrr::map(function(x)
    select(x, starts_with("sim")) %>%
      mRIIDS:::shuffle_cols() %>%
    colSums())

aggregated <- purrr::map_dfr(
  aggregated, function(x)
    data.frame(matrix(x, nrow = 1, dimnames = list("1", names(x)))),
  .id = "date"
)



aggregated$date <- as.Date(aggregated$date)
wafrica_weekly <- mRIIDS:::daily.to.weekly(aggregated)

wafrica_distr <- tidyr::gather(wafrica_weekly,
                               sim,
                               aggregated,
                               -date) %>%
    mRIIDS:::projection_quantiles()

projections_distr <- rbind(projections_distr, wafrica_distr)

```


And finally write the them out for later validation and testing.

```{r write}
outfile <- paste0("forecasts_",
                  params$tproj,
                  "_",
                  params$twindow,
                  "_",
                  params$n.dates.sim,
                  ".csv")
here::here(params$outdir,
           outfile) %>%
    readr::write_csv(x = projections_distr,
                     path = .)

```


Also save all samples.


```{r writeall}
outfile <- paste0("forecasts_samples_",
                  params$tproj,
                  "_",
                  params$twindow,
                  "_",
                  params$n.dates.sim,
                  ".Rds")
here::here(params$outdir,
           outfile) %>%
    readr::write_rds(x = daily_projections,
                     path = .)

```
