---
output: 
pdf_document:
citation_package: natbib
keep_tex: false
fig_caption: true
latex_engine: pdflatex
title: Reproduction number as a function of the time window 
author:
- name: Sangeeta Bhatia
affiliation: Imperial College London
abstract: 
keywords: 
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  tproj: 84
  twindow: !r c(7, 14, 21, 28, 42)
  N: 3
  incid: healthmap_wide.csv
---

```{r setup, eval = TRUE}
library(dplyr)
library(purrr)
library(ggplot2)
library(ggthemes)
```

## Read incidence data

```{r incid}
incid <- here::here("data/processed", params$incid) %>%
    readr::read_csv(n_max = params$tproj)
countries <- colnames(select(incid, -date))
```


The shorter the time window, the more the number of parameters in our
model. We use the reproduction numbers in the last window to project.
Hence we will look at how the estimates vary with the time window for
a given point at which we are projection.

```{r fits}
suffix <- paste0(params$tproj, "_", params$twindow, ".rds")
names(suffix) <- paste0("at_", params$twindow)
fits <- map(suffix,
            ~ readRDS(here::here("data/stanfits",
                                 .x)))

rsamples <- map(fits, ~ rstan::extract(.x)[["R"]])
```
In each case, the last N variables are of particular interest as 
these are the ones that we use for projection.

```{r}
rsamples <- map_dfr(rsamples, function(x) {
    df <- x[ , (ncol(x) - params$N + 1 ):ncol(x)]
    colnames(df) <- countries
    as.data.frame(df)
}, .id = "at")

rsamples <- tidyr::gather(rsamples, country, R, -at)
```

```{r viz}
rsamples$at <- factor(rsamples$at,
                      levels = names(suffix))
p <- ggplot(rsamples, aes(at,
                     R)) +
    geom_point(alpha = 0.1) +
    facet_wrap(~country, nrow = 3, scales = "free_y") +
    theme_classic()
```

```{r save}
outfile <- here::here("data/output",
                      paste0("r_vs_twindow_", params$tproj, ".png"))

ggpubr::ggarrange(p, align = "hv", ncol = 1) %>%
    ggpubr::ggexport(filename = outfile)
```
