---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
title: Comparison of incidence data from ProMED, HealthMap and WHO
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "`r Sys.Date()`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
---


```{r who-hm-promed-comparison-1, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dev = "pdf",
  dpi = 300,
  fig.width = mriids_plot_theme$double_col_width / 2.5,
  fig.height = mriids_plot_theme$double_col_width / 5,
  ## dev.args = list(pointsize = 6),
  fig.path = "figures/"
)
```

```{r setup, echo = FALSE}
library(ggplot2)
library(ggthemes)
library(dplyr)
library(EpiEstim)
library(purrr)
library(ggpubr)
```



## Comparison of daily incidence trends


```{r who-hm-promed-comparison-2, echo = FALSE}

datasources <- c("ProMED", "HealthMap", "WHO")
names(datasources) <- datasources
daily_incid <- purrr::map_dfr(
  datasources,
  function(datasource) {
    infile <- here::here(
      all_files[[datasource]]$incidtall
    )
    message("Reading ", infile)
    out <- readr::read_csv(infile)
    out
  },
  .id = "datasource"
)
daily_incid <- daily_incid[daily_incid$country %in%
  c("LBR", "GIN", "SLE"), ]
```


```{r who-hm-promed-comparison-3, eval = TRUE}

daily_incid$country <- forcats::fct_recode(
  daily_incid$country,
  Guinea = "GIN",
  Liberia = "LBR",
  `Sierra Leone` = "SLE"
)

## Comparison of daily incidence trends
p_daily <-
  ggplot(
    daily_incid,
    aes(date,
      incid,
      color = interaction(datasource, interpolated)
    )
  ) +
  geom_point(shape = 20, shape = 1)

p_daily <- p_daily +
  mriids_plot_theme$theme

p_daily <- p_daily +
  facet_wrap(country ~ .,
    scales = "free_y",
    strip.position = "top",
    nrow = 3
  ) +
  scale_colour_manual(
    values = mriids_plot_theme$color_scale,
    breaks = c(
      "ProMED",
      "HealthMap",
      "WHO"
    )
  )

p_daily <- p_daily +
  mriids_plot_theme$legend

p_daily <- p_daily +
  xlab("") +
  ylab("Daily Incidence")

p_daily <- p_daily +
  scale_x_date(labels = mriids_plot_theme$dateformat) +
  mriids_plot_theme$xticklabels
# p_daily <- p_daily +
#   ##     theme(axis.text.x = element_text(angle = -10, hjust = 0)) +
#   scale_x_date(
#     date_breaks = "1 year",
#     labels = scales::date_format("%b %Y")
#   )

#
# p_daily <- p_daily +
#   theme(legend.title = element_blank())

p_daily
```


## Estimating Reproduction number from WHO and Healthmap Data

# Parameters for Ebola

Culled from literature.

```{r who-hm-promed-comparison-4, eval = TRUE}

mean_SI <- 14.2
CV_SI <- 9.6 / 14.2
SItrunc <- 40
SI_Distr <- sapply(
  0:SItrunc,
  function(e) EpiEstim::DiscrSI(e, mean_SI, mean_SI * CV_SI)
)
SI_Distr <- SI_Distr / sum(SI_Distr)
```
R estimate is strongly influenced by the choice of time window and we
want to capture that. Hence, estimate R for different values of time
window.

```{r who-hm-promed-comparison-5 }

all_groups <- split(
  daily_incid,
  list(
    daily_incid$datasource,
    daily_incid$country
  )
)

all_R <- purrr::map_dfr(
  list(`2` = 2, `4` = 4, `6` = 6),
  function(tw) {
    time_window <- 7 * tw
    df <- purrr::map_dfr(
      all_groups,
      function(x) {
        I <- x$incid
        start <- 2:(nrow(x) - time_window)
        end <- start + time_window
        res <- EpiEstim::EstimateR(
          I,
          T.Start = start,
          T.End = end,
          method = "NonParametricSI",
          SI.Distr = SI_Distr,
          plot = FALSE,
          CV.Posterior = 1,
          Mean.Prior = 1,
          Std.Prior = 0.5
        )
        res$R$date <- x$date[end]
        ## Append an interpolated = TRUE flag if
        ## any point in the window was interpolated.
        res$R$interpolated <- purrr::map2_lgl(
          start,
          end,
          function(st, ed) {
            all(x$interpolated[st:ed])
          }
        )
        res$R
      },
      .id = "source.country"
    )
    df
  },
  .id = "time_window_in_weeks"
)


all_R <- tidyr::separate(
  all_R,
  source.country,
  into = c("Source", "country"),
  extra = "merge",
  fill = "left"
)
```   

Checking how the choice of time window impacts R estimates and
correlation.

```{r who-hm-promed-comparison-6 }

all_median_R <- dplyr::select(
  all_R,
  time_window_in_weeks,
  Source,
  country,
  `Median(R)`,
  date
)


all_median_R <- tidyr::spread(
  all_median_R,
  Source,
  `Median(R)`
)

## dim(all_median_R)
## [1] 6669    6
## > dim(na.omit(all_median_R))
## [1] 5643    6

all_median_R <- na.omit(all_median_R)
```

Correlation between R series.

```{r who-hm-promed-comparison-7 }
byparams <- split(
  all_median_R,
  list(
    all_median_R$country,
    all_median_R$time_window_in_weeks
  ),
  sep = "_"
)

coeffs <- purrr::map_dfr(
  byparams,
  function(df) {
    df <- dplyr::select(df, WHO, ProMED, HealthMap)
    out <- Hmisc::rcorr(as.matrix(df), type = "pearson")
    out <- broom::tidy(out)
    out
  },
  .id = "params"
)

coeffs <- tidyr::separate(
  coeffs,
  col = "params",
  into = c("country", "time_window_in_weeks"),
  sep = "_"
)

coeffs <- dplyr::mutate_if(
  coeffs, is.numeric, round, 2
)

knitr::kable(coeffs)
```

```{r who-hm-promed-comparison-8}
## Impact of time_window on R estimation and correlation
pm <- dplyr::filter(coeffs, column1 == "WHO", column2 == "ProMED")
hm <- dplyr::filter(coeffs, column1 == "WHO", column2 == "HealthMap")
pm_hm <- dplyr::filter(coeffs, column1 == "ProMED", column2 == "HealthMap")
pm$label <- glue::glue("R(WHO-ProMED): {pm$estimate}")
hm$label <- glue::glue("R(WHO-HealthMap): {hm$estimate}")
pm_hm$label <- glue::glue("R(ProMED-HealthMap): {pm_hm$estimate}")
p_r_tw <- ggplot(all_median_R) +
  geom_point(aes(
    x = WHO,
    y = ProMED
  ),
  col = mriids_plot_theme$color_scale[["ProMED"]],
  shape = 20, size = 0.5
  ) +

  geom_point(aes(
    x = WHO,
    y = HealthMap
  ),
  col = mriids_plot_theme$color_scale[["HealthMap"]],
  shape = 20, size = 0.5
  ) +
  xlim(0, 7) + ylim(0, 7) +
  geom_abline(
    slope = 1, intercept = 0, alpha = 0.2, linetype = "dashed"
  ) +
  geom_text(
    data = pm, aes(x = 4, y = 1, label = label),
    color = mriids_plot_theme$color_scale[["ProMED"]],
    size = 5 / .pt
  ) +
  geom_text(
    data = hm, aes(x = 4, y = 1.7, label = label),
    color = mriids_plot_theme$color_scale[["HealthMap"]],
    size = 5 / .pt
  ) +
  geom_text(
    data = pm_hm, aes(x = 4, y = 2.5, label = label),
    size = 5 / .pt
  ) +

  facet_grid(time_window_in_weeks ~ country,
    scales = "fixed",
    labeller = labeller(
      time_window_in_weeks =
        as_labeller(c(
          `2 weeks` = 2,
          `6 weeks` = 6
        ))
    )
  )

p_r_tw <- p_r_tw +
  mriids_plot_theme$theme

p_r_tw <- p_r_tw +
  xlab("Median(R) WHO") +
  ylab("Median(R) ProMED/HealthMap")

p_r_tw
```









```{r who-hm-promed-comparison-9 }
all_R_6 <- dplyr::filter(all_R, time_window_in_weeks == 4)
p <- ggplot() +
  mriids_plot_theme$theme +
  mriids_plot_theme$legend

p <- p + geom_ribbon(
  data = all_R_6,
  aes(
    x = date,
    ymin = `Quantile.0.025(R)`,
    ymax = `Quantile.0.975(R)`,
    fill = Source
  ),
  alpha = 0.3
)
p <- p + geom_line(
  data = all_R_6,
  aes(
    x = date,
    y = `Median(R)`,
    color = Source
  )
)

p <- p + facet_wrap(country ~ .,
  scale = "fixed",
  strip.position = "top",
  nrow = 3
)

p <- p + xlab("") + ylab("R")
p <- p +
  scale_colour_manual(
    values = mriids_plot_theme$color_scale,
    breaks = c(
      "ProMED",
      "HealthMap",
      "WHO"
    )
  ) +
  scale_fill_manual(
    values = mriids_plot_theme$color_scale,
    breaks = c(
      "ProMED",
      "HealthMap",
      "WHO"
    )
  )

p <- p +
  geom_hline(
    yintercept = 1,
    linetype = "dashed"
  )

p <- p + scale_x_date(labels = mriids_plot_theme$dateformat) +
  mriids_plot_theme$xticklabels
## Add correlation coefficient
## median_coeffs <- mutate_if(
##   median_coeffs,
##   is.numeric,
##   ~ round(.x, digits = 2)
## )

## p2 <- p +
##   geom_text(
##     data = median_coeffs,
##     aes(
##       x = as.Date("2015-07-06"),
##       y = 4,
##       label =
##         paste0(
##           "   W  H  P \n",
##           "W \n",
##           "H ", WHO_HealthMap, "\n",
##           "P ", WHO_ProMED, " ", HealthMap_ProMED
##         )
##     ),
##     lineheight = .7,
##     size = 2
##   )


p
```




```{r who-hm-promed-comparison-10 }

p_all <- ggarrange(p_daily,
  p,
  nrow = 1,
  ncol = 2,
  labels = "AUTO",
  legend = NULL
)

p_all

ggplot2::ggsave(
  filename = here::here(
    "ms-figures",
    glue::glue("{Sys.Date()}_who-hm-promed-comparison.pdf")
  ),
  plot = p_all,
  width = mriids_plot_theme$double_col_width,
  height = mriids_plot_theme$double_col_height,
  units = mriids_plot_theme$units
)

## ggexport(p_all,
##   filename = here::here(
##     "data/output/figures",
##     "who_hm_promed_compare.eps"
##     ),
##   width = 11.4,
##   units = "cm"
## )
```




```{r who-hm-promed-comparison-11 }

readr::write_csv(
  x = all_R,
  path = here::here("data/output", "Rt_who_pm_hm.csv")
)

readr::write_csv(
  x = coeffs,
  path = here::here(
    "data/output",
    "Rt_who_pm_hm_correlation.csv"
  )
)
```
```{r who-hm-promed-comparison-12 }
devtools::session_info()
```
