---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
title: WHO Data Pre-processing
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "Y"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=6, echo=FALSE, warning=FALSE, message=FALSE, fig.path = "figures/")
```

```{r setup, echo = FALSE}
library(magrittr)
library(ggplot2)
library(ggthemes)
library(dplyr)
devtools::load_all()

```

# Introduction

The pre-processing steps consist of turning the line count into a
daily incidence series and turning the given tall data into wide data.


The current data set has three possible values for the epidemiological
case definition - confirmed, probable and suspected.
Epidemiological case definition of probable and suspected cases
differed across countries while confirmed cases were the ones
confirmed through lab report. For the purpose of the current analysis,
one approach could be to lump all of them together. The inferred date of
onset (rather than the date reported) is used for estimation.
The columns we use are : Country, EpiCaseDef (probably),
DateOnsetInferred and CL_DistrictRes. 


```{r rawdata, eval = TRUE}


who_raw <- here::here("data/raw", "rstb20160308supp1.csv") %>%
           read.csv(colClasses = c(Country = "factor",
                                   EpiCaseDef = "character",
                                   DateOnsetInferred = "Date",
                                   CL_DistrictRes = "factor")) %>%
           select(Country, EpiCaseDef,
                  DateOnsetInferred,CL_DistrictRes) 
```

## Line-list to incidence time series

This can be easily accomplished using the `incidence` package.

```{r ll}
incid <- split(who_raw, who_raw$Country) %>%
    map_dfr(~ incidence::incidence(.x$DateOnsetInferred) %>%
                as.data.frame,
            .id = "country")


incid <- tidyr::spread(incid, country, counts, fill = 0)
incid <- rename(incid, "date" = "dates")
```

```{r save}
outfile <- here::here("data/processed", "who_bycountry.csv")
readr::write_csv(x = incid, path = outfile)
```
