---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
title: Comparison of incidence data from ProMED, HealthMap and WHO
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "`r Sys.Date()`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  ofinterest: !r c("GIN", "LBR", "SLE")
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE, fig.path = "figures/")
```

```{r setup, echo = FALSE}
library(ggplot2)
library(ggthemes)
library(dplyr)
library(EpiEstim)
library(purrr)

```

# Parameters for Ebola

Culled from literature.

```{r ebola_params, eval = TRUE}

mean_SI     <- 14.2
CV_SI       <- 9.6 / 14.2
SItrunc     <- 40
SI_Distr    <- sapply(0:SItrunc, function(e) EpiEstim::DiscrSI(e, mean_SI, mean_SI * CV_SI))
SI_Distr    <- SI_Distr / sum(SI_Distr)
time_window <- 7 * 2

```


First read in the cleaned up data from Health Map. Since these data
are at national level, we will also aggregate the WHO data at country
level. HealthMap collects and curates data on the cumulative number
of cases (C) and deaths (D) stratified by status (suspected, SC/SD, or
confirmed CC/CD). For data from the WHO, the case status could be
confirmed, probable or suspected. The final clinical outcome (alive/dead)
for each case is also recorded. In comparing the two data sets, we sum
across all four categories in Healthmap and derive incidence data from
cumulative data.

```{r hmread, echo = FALSE}
healthmap <- here::here("data/processed",
                        "04022019_healthmap_loglinear_wide.csv") %>%
    readr::read_csv() %>%
    select(date, params$ofinterest)



hm_weekly <- mRIIDS:::daily.to.weekly(healthmap)

hm_weekly$date <- lubridate::ymd(hm_weekly$date)

```

ProMED data is similarly stratfied and we will sum across the
categories of suspected and confirmed cases for ProMED.

```{r pmread, echo=FALSE}

promed <- here::here("data/processed",
                     "20122018_promed_loglinear_wide.csv") %>%
    readr::read_csv() %>%
    select(date, params$ofinterest)    


promed_weekly <- mRIIDS:::daily.to.weekly(promed)

promed_weekly$date <- as.Date(promed_weekly$date)

```


```{r whoread, eval = TRUE}

who_bycountry <- here::here("data/processed",
                            "who_bycountry.csv") %>%
    readr::read_csv() 
    
colnames(who_bycountry) <- c("date",
                             "GIN",
                             "LBR",
                             "SLE")
who_weekly <- mRIIDS:::daily.to.weekly(who_bycountry)   
who_weekly$date <- as.Date(who_weekly$date)

```


```{r weekly_incid_viz, eval = TRUE}

color_scale <- c("HealthMap" = "black",
                 "WHO" = "green",
                 "ProMED" = "red")

all_weekly <- list(WHO = who_weekly,
                   HealthMap = hm_weekly,
                   ProMED    = promed_weekly) %>%
    bind_rows(.id = "Source")

all_weekly <- tidyr::gather(all_weekly,
                            country,
                            incid,
                            - c(Source:date))
p <- 
    ggplot(all_weekly,
           aes(date, incid, color = Source)) +
    geom_point() +
    facet_grid(country ~., scales = "free_y") +
    scale_colour_manual(values = color_scale)

p <- p + theme(legend.title = element_blank())
p <- p + xlab("") + ylab("Weekly Incidence")
p <- p + theme_classic()

p

```

## Correlation between the incidence data

```{r incid_corrplot, eval = TRUE}
library(zeallot)
c(who, pm, hm) %<-%  map(list(who_weekly,
                              promed_weekly,
                              hm_weekly),
                         ~ tidyr::gather(.x,
                                         country,
                                         incid,
                                         -date))

c(who$week, pm$week, hm$week) %<-%
    map(list(who, pm, hm),
        ~ paste(lubridate::year(.x$date),
                lubridate::week(.x$date),
                sep = "-"))

hm_who_pm_incid <- inner_join(who, hm,
                              by = c("week", "country")) %>%
              select(country, week, incid.x, incid.y) %>%
              rename("WHO" = `incid.x`,
                     "HealthMap" = `incid.y`) %>%
              inner_join(pm, by = c("week", "country")) %>%
              rename("ProMED" = `incid`) %>%
              select(WHO, ProMED, HealthMap)


M <- Hmisc::rcorr(as.matrix(hm_who_pm_incid), type = "pearson")

```

## Estimating Reproduction number from WHO and Healthmap Data


```{r restim}
all_R <- split(all_weekly, list(all_weekly$Source,
                                all_weekly$country)) %>%
    map_dfr(function(x) {
        I <- x$incid
        start <- 2:(nrow(x) - time_window)
        end <- start + time_window
        res <- EpiEstim::EstimateR(I,
                                   T.Start = start,
                                   T.End = end,
                                   method   = "NonParametricSI",
                                   SI.Distr = SI_Distr,
                                   plot     = FALSE ,
                                   CV.Posterior = 1 ,
                                   Mean.Prior   = 1 ,
                                   Std.Prior    = 0.5)
        res$R$date <- x$date[end]
        res$R
                                   
    },
    .id = "source.country")


all_R <- tidyr::separate(all_R,
                         source.country,
                         into = c("Source", "country"),
                         extra = "merge",
                         fill = "left")


```   


```{r restimviz}

p <- ggplot()
p <- p + geom_ribbon(data = all_R,
                     aes(x = date,
                         ymin = `Quantile.0.025(R)`,
                         ymax = `Quantile.0.975(R)`,
                         fill = Source),
                     alpha = 0.3)
p <- p + geom_line(data = all_R,
                   aes(
                       x = date,
                       y = `Mean(R)`,
                       color = Source))
p <- p + facet_grid(country ~., scale = "free_y")
p <- p + theme(legend.title = element_blank())
p <- p + xlab("")
p <- p +
    scale_colour_manual(values = color_scale) +
    scale_fill_manual(values = color_scale) 
p <- p + ggtitle("Mean R estimated from incidence data") +
         theme(plot.title = element_text(hjust = 0.5))
p <- p + theme_classic()
p

```

The graphs show that the estimates from the three different sources
closely track each other. Here's another way to visualise this.
As the above graphs indicates, estimates of the reproduction number
from ProMed, HealthMap and WHO data are highly correlated (correlation
coefficient = 0.76, 95% CI = (0.74, 0.78)).

