---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
title: WHO Data Pre-processing
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "Y"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  infile: data/raw/rstb20160308supp1.csv
  outfile: data/processed/15032019_who_bycountry.csv
  ofinterest: data/processed/all_african_centroids.csv
---
```{r who-data-processing-1, include=FALSE}
knitr::opts_chunk$set(fig.width=12,
                      fig.height=6,
                      echo=FALSE,
                      warning=FALSE,
                      message=FALSE,
                      fig.path = "figures/")
```

```{r setup, echo = FALSE}
library(ggplot2)
library(ggthemes)
library(dplyr)
library(purrr)

```

# Introduction

The pre-processing steps consist of turning the line count into a
daily incidence series and turning the given tall data into wide data.


The current data set has three possible values for the epidemiological
case definition - confirmed, probable and suspected.
Epidemiological case definition of probable and suspected cases
differed across countries while confirmed cases were the ones
confirmed through lab report. For the purpose of the current analysis,
one approach could be to lump all of them together. The inferred date of
onset (rather than the date reported) is used for estimation.
The columns we use are : Country, EpiCaseDef (probably),
DateOnsetInferred and CL_DistrictRes. 


```{r who-data-processing-2, eval = TRUE}


who_raw <- here::here(params$infile) %>%
           read.csv(colClasses = c(Country = "factor",
                                   EpiCaseDef = "character",
                                   DateOnsetInferred = "Date",
                                   CL_DistrictRes = "factor")) %>%
           select(Country, EpiCaseDef,
                  DateOnsetInferred,CL_DistrictRes) 
```

## Line-list to incidence time series

This can be easily accomplished using the `incidence` package.

```{r who-data-processing-3 }
incid <- split(who_raw, who_raw$Country) %>%
    map_dfr(~ incidence::incidence(.x$DateOnsetInferred) %>%
                as.data.frame,
            .id = "country")


incid <- tidyr::spread(incid, country, counts, fill = 0)
incid <- rename(incid, "date" = "dates")
```


## Add 0 incidence for other countries

Finally, we add 0 incidence for all countries other than ones for
which we already have data. 

```{r who-data-processing-4 }

ofinterest <- readr::read_csv(here::here(params$ofinterest)) %>%
    filter(include == "TRUE")

wafrica <- select(incid, -date) %>% colnames
colnames(incid)[-1] <-  countrycode::countrycode(wafrica,
                                                 'country.name',
                                                 'iso3c')
nodata <- setdiff(ofinterest$ISO3,
                  colnames(incid)[-1])
df <- matrix(0,
             nrow = nrow(incid),
             ncol = length(nodata))

colnames(df) <- nodata
df <- data.frame(df)
incid_wide_all <- cbind(incid, df)

```

Finally ensure that the order of countries in centroids and incid 
files is the same.

```{r who-data-processing-5, eval = TRUE}
inorder <- ofinterest$ISO3
## Some of the countries of interest may not be present in
## incid_wide_all
inorder <- intersect(inorder, colnames(incid_wide_all))
incid_wide_all <- select(incid_wide_all,
                         date,
                         inorder)
```

```{r who-data-processing-6 }
outfile <- here::here(params$outfile)
readr::write_csv(x = incid_wide_all,
                 path = outfile)
```


Other data files we will need later.

```{r}

## Weekly Trend
who_weekly <- mRIIDS:::daily.to.weekly(incid)
who_weekly <- tidyr::gather(
  who_weekly,
  country,
  incid,
  -date
)
who_weekly$interpolated <- FALSE
who_weekly$date <- lubridate::ymd(who_weekly$date)

outfile <- stringr::str_replace(params$outfile,
                                ".csv",
                                "_weekly.csv")

readr::write_csv(x = who_weekly,
                 path = here::here(outfile))
```

```{r}
## Cumulative
who_cumulative <- dplyr::mutate_if(
  incid,
  is.numeric,
  cumsum
) %>%
  tidyr::gather(country, cases, -date) %>%
dplyr::select(country, date, cases)

outfile <- stringr::str_replace(params$outfile,
                                ".csv",
                                "_cumulative.csv")

readr::write_csv(x = who_cumulative,
                 path = here::here(outfile))
```

```{r}
## For stacking with ProMED and HM data
## that we will read in tall format
who_tall <- tidyr::gather(
  incid,
  country,
  incid,
  -date
)
who_tall$interpolated <- FALSE

outfile <- stringr::str_replace(params$outfile,
                                ".csv",
                                "_tall.csv")

readr::write_csv(x = who_tall,
                 path = here::here(outfile))

```
